---
title: "Orang Asli Resistome"
author: "Li-Fang Yeo"
format: html
editor: visual
---

### Resistome analysis

The study of the gut resistome of indigenous Orang Asli communities in Peninsular Malaysia, Temuan (urban), Temiar (semi-rural) and Jahai (hunter-gatherer). We used an urban Malay community as a comparator group.

```{r,include=FALSE}
#load libraries
library(mia)
library(tidyverse)
library(dplyr)
library(knitr)
library(miaViz)
library(stringr)
library(maaslin3)
library(tableone)
library(ggsignif)
```

# Resfinder

ARGs were mapped to Resfinder database

```{r}
#Data input 
#file containting RPKM per ARG from resfinder
kma_out <- read_csv("resistome/resistome/resfinder/kma_out.csv")

#files needed to build tse
counts <- kma_out %>% 
  dplyr::select(sampleID, refSequence, RPKM) %>%
  tidyr::pivot_wider(names_from = sampleID, values_from = RPKM)

#this is the whole resfinder database
tax <- read.csv("C:/Users/lifyeo/oa-microbiome/resistome/resistome/resfinder/raw/resfinder_db.csv")

tax_list <- counts %>% 
  dplyr::select(refSequence)

#to match the number of tax to abundance table  
tax <- tax_list %>% dplyr::left_join(tax, join_by("refSequence"))
tax <- tax %>% tibble::column_to_rownames(var = "refSequence")
counts <- counts %>% tibble::column_to_rownames("refSequence")
meta <- read.csv("Metafile.csv", row.names = 1, header = TRUE)

##QC to check data format and matching rows
#match rows and columns
genomic_idx <- match(rownames(meta), colnames(counts))
genomic_idx
counts_reordered  <- counts[ , genomic_idx]
all(rownames(meta) == colnames(counts_reordered))
all(rownames(tax) == rownames(counts_reordered))
#data needs to be in numeric matrix format
counts_reordered <- as.matrix(counts_reordered)
# Replace NAs with zeroes
counts_reordered[is.na(counts_reordered)] <- 0
# Remove rows and cols with zero variance
counts_reordered <- counts_reordered[rowSds(counts_reordered)>0, colSds(counts_reordered)>0]

class(meta)
class(tax)
class(counts_reordered)

tse_arg <- TreeSummarizedExperiment(
    assays =  SimpleList(counts = counts_reordered),
    colData = DataFrame(meta),
    rowData = DataFrame(tax))

tse_arg
```

Alpha diversity - Shannon

```{r}
tse_arg <- mia::addAlpha(tse_arg, 
                              assay.type = "counts",
                              index = "shannon", 
                              name = "shannon")
df<- as.data.frame(colData(tse_arg))

#change level
df$Group <- factor(df$Group, levels = c("Jahai", "Temiar", "Temuan", "Malay"))
# For significance testing, all different combinations are determined
comb <- split(t(combn(levels(df$Group), 2)), 
           seq(nrow(t(combn(levels(df$Group), 2)))))
#coloured by increasing rural-ness

my_colors <- c("Malay"="darkkhaki", "Temuan"="#8eb79c","Temiar"= "#4b705c","Jahai" = "black")
#pdf("alpha_arg.pdf")
p<- ggplot(df, aes(x = Group, y = shannon, fill = Group)) +
  # Outliers are removed, because otherwise each 
  # data point would be plotted twice; 
  # as an outlier of boxplot and as a point of dotplot.
  geom_boxplot(outlier.shape = NA) + 
  geom_jitter(width = 0.2) + 
  geom_signif(comparisons = comb, map_signif_level = FALSE,
               step_increase = 0.1) + #Corrects the p-values and step increase minimise overlap
  theme(text = element_text(size = 6)) + theme_classic() +
  scale_fill_manual(values = my_colors)

ggsave("alpha_arg.pdf",
       p, 
       width = 8,
       height = 6,
       units = "in")

```

Beta diversity - Bray curtis

```{r}
x1 <- runRDA(tse_arg,
               assay.type = "counts",
               formula = assay ~ Group + Reads,
               distance = "bray",
               na.action = na.exclude)

# Store results of PERMANOVA test
rda_info <- attr(reducedDim(x1, "RDA"), "significance")
```

PERMANOVA

```{r}
rda_info$permanova |>
  knitr::kable()

write.csv(rda_info$permanova, "permanova_ARG.csv")
```

```{r}
#homogeneity test
rda_info$homogeneity |>
  knitr::kable()

write.csv(rda_info$homogeneity, "homogeneity_ARG.csv")
```

RDA plot

```{r}
#dbRDA plot
p <- plotRDA(x1, "RDA", add.ellipse = TRUE, color_by = "Group") +
  scale_color_manual(values = my_colors, name = "Groups",
  breaks = c("Jahai", "Temiar", "Temuan", "Malay")) +
  theme_classic() 
p
```

```{r}
#plot dbRDA loadings
L <- miaViz::plotLoadings(x1, "RDA", ncomponents = 2, n = 10)

L<- L + scale_fill_manual(name = "Effect", 
                          values = c("lightblue", "maroon"),
                          labels = c('Negative', 'Positive'))

L
library(patchwork)
p / L
onni <- p / L
ggsave("RDA-loadings-ARG.pdf",
       onni, 
       width = 8,
       height = 8,
       units = "in")
```
Differential abundance analysis - resfinder
```{r}
tse4 <- tse_arg %>% 
  mia::agglomerateByPrevalence(assay.type = "counts",
                               detection =  0.1/100,
                               prevalence = 5/100)
  

maaslin3_out <- maaslin3(
  input_data = tse4,
  output = "ARG_DAA1",
  formula = " ~ Group + Reads",
  normalization = "CLR",
  transform = "NONE",
  verbosity = "ERROR",
  standardize = "TRUE",
  reference = "Group,Malay",
  warn_prevalence = "FALSE",
  max_significance = 0.05,
  correction = "BH",
  plot_summary_plot = TRUE,
  plot_associations = FALSE
)


```


